---
title: "review_GLMM_glmmTMB_v2"
date: "2025-12-05"
output: pdf_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

```{r}
.libPaths(c(.libPaths(),"/dss/dsshome1/lxc06/ru87mis/Rstudio/4.4"))

#set wd
setwd('/dss/dsslegfs01/pr53da/pr53da-dss-0029/big_assortative')

#Load the file. 
hoppers <- read.csv("hoppers2_femresp_corrected.csv", sep = ";")
#View(hoppers)


## THIS VERSION USES THE PACKAGE GLMMTMB INSTEAD OF LME4 TO FIT THE GLMMs

# Install required packages 
install.packages("glmmTMB") #GLMM modeling
install.packages("bbmle") #required for glmmTMB package
install.packages('DHARMa') #goodness of fit test
install.packages('emmeans') #strength of RI
install.packages('car')

# Load packages
library(dplyr)
library(stringr)
#library(reshape)
#library(reshape2) 
library(tidyr)
#library(lme4)
library(ggplot2)
library(car)

library(glmmTMB)
library(bbmle)
library(DHARMa)
library(emmeans)
```

# REVIEW 1

Date of review: Oct 2025 Date of edits: Dec 2025

Here, we try to address the issues raised by the reviewers for the first version of our submitted paper.

Changes:

\- After discussing with Dirk, we will add the origin (Type.F/Type.M) to the model as a fixed effect factor.

\- To avoid confusion, for the reviewed version, we have changed the naming of the models.
The reduced models are now called additive models and the saturated models are not called interaction models.

\- To check whether the results are not affected by the type of analysis, i.e.
GLM or GLMMM, I have run the GLMs for all behaviors and combinations, not only for those that raised convergence or singularity issues.
-- This section is no longer relevant, since only a single model raised convergence issues with the glmmTMB pacakge.

\- To improve model fitting and reduce singularity and convergence issues, we used the 'glmmTMB' package, which did not throw back singularity or convergence issues, and therefore we will move forward with this package for our analysis.
- Check model fit and residual plots using DHARMa package.

\- Since DHARMa showed issues for most of the models in one way or another, here, in v2 of the code, first, we are replacing the family type for fitting the models from poisson to negative binomial (nbinom2), for those that used a poisson distribution.

Species Key:

Bru - C.
brunneus

Big - C.biguttulus

Mol - C.mollis

# DATA CURATION

For the Female Behaviors we need to use a binomial distribution, and use data on the success and fail events on each experiment.

```{r}
#Creating a fail column to convert into sucess/fail data 
hoppers$Response_fail <- hoppers$Total.Singing - hoppers$Response
hoppers$Rejection_fail <- hoppers$Harassment - hoppers$Hrejection

```

For the female behaviors we created separate datasets so we can delete the rows where males didn't perform the behavior.

```{r}
# Ensure origin variables are factors
hoppers$Type.F <- factor(hoppers$Type.F)  # lab / wild
hoppers$Type.M <- factor(hoppers$Type.M)

# Flags and combined origin
hoppers$FemaleWild <- hoppers$Type.F == "wild"
hoppers$MaleWild   <- hoppers$Type.M == "wild"
hoppers$AnyWild    <- hoppers$FemaleWild | hoppers$MaleWild
hoppers$BothLab    <- !hoppers$FemaleWild & !hoppers$MaleWild

hoppers$OriginPair <- interaction(hoppers$Type.F, hoppers$Type.M, sep = "_")

#Response 
hoppers$ResponseNorm <- (hoppers$Response/hoppers$Total.Singing)  
resp <- hoppers[-which(hoppers$ResponseNorm == "NaN"), ]      
#resp <- resp[-which(resp$ResponseNorm == "Inf"), ]

#Rejection
hoppers$RejectNorm <- (hoppers$Hrejection/hoppers$Harassment)
rej <- hoppers[-which(hoppers$RejectNorm == "NaN"), ] 

# Response: remove rows where Total.Singing = 0 (NaN in proportion)
resp <- subset(hoppers, !is.nan(ResponseNorm))

# Rejection: remove rows where Harassment = 0 (NaN in proportion)
rej  <- subset(hoppers, !is.nan(RejectNorm))
```

To run the GLMM we need to subset the data for the three possible combinations of species.
First, we work with male behaviors.
These are the subsets for the male behaviors

```{r}
lm_brubig <- hoppers %>% filter(Combination == "BruBru" | Combination == "BruBig"|
                                Combination == "BigBru" | Combination == "BigBig")

lm_brumol <- hoppers %>% filter(Combination == "BruBru" | Combination == "BruMol" |
                                Combination == "MolBru" | Combination == "MolMol")

lm_molbig <- hoppers %>% filter(Combination == "BigBig" | Combination == "BigMol" |
                                  Combination == "MolBig" | Combination == "MolMol"
```

# MODEL ANALYSES

Naming as: [model:additive/interaction]\_[combination]

Male behaviors per combination

## Combination: Bru - Big

```{r}
                            ### Combination: Bru - Big ###

# Behavior: Call Singing (Indirect Singing) 
interaction_indirect_brubig <- glmmTMB(Singing ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                            family = nbinom2, data = lm_brubig) #m1


additive_indirect_brubig <- glmmTMB(Singing ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                               family = nbinom2, data = lm_brubig)

anova(interaction_indirect_brubig, additive_indirect_brubig)
```

```{r}
# Behavior: Courtship singing (Direct Singing) 
interaction_direct_brubig <- glmmTMB(Active.Singing ~ Male*Female + Type.F + Type.M +
(1|Cage.males) + (1|Cage.females), 
                            family = nbinom2, data = lm_brubig)

additive_direct_brubig <- glmmTMB(Active.Singing ~ Male + Female + Type.F + Type.M + 
(1|Cage.males) + (1|Cage.females),
                               family = nbinom2, data = lm_brubig)

anova(interaction_direct_brubig, additive_direct_brubig)

```

```{r}
# Behavior: Mounting attempt (Harassment)
interaction_harass_brubig <- glmmTMB(Harassment ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                            family = nbinom2, data = lm_brubig)

additive_harass_brubig <- glmmTMB(Harassment ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                               family = nbinom2, data = lm_brubig)

anova(interaction_harass_brubig, additive_harass_brubig)
```

## Combination: Bru - Mol

```{r}
                            ### Combination: Bru - Mol ### 

#Behavior: Call Singing (Indirect Singing)
interaction_indirect_brumol <- glmmTMB(Singing ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                            family = nbinom2, data = lm_brumol)


additive_indirect_brumol <- glmmTMB(Singing ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                               family = nbinom2, data = lm_brumol)

anova(interaction_indirect_brumol, additive_indirect_brumol)
```

```{r}
# Behavior: Courtship singing (Direct/Active Singing)
interaction_direct_brumol <- glmmTMB(Active.Singing ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                            family = nbinom2, data = lm_brumol)

additive_direct_brumol <- glmmTMB(Active.Singing ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                               family = nbinom2, data = lm_brumol)

anova(interaction_direct_brumol, additive_direct_brumol)

```

```{r}
# Behavior: Mounting attempt (Harassment)
interaction_harass_brumol <- glmmTMB(Harassment ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                            family = nbinom2, data = lm_brumol)

additive_harass_brumol <- glmmTMB(Harassment ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                               family = nbinom2, data = lm_brumol)

anova(interaction_harass_brumol, additive_harass_brumol)

```

## Combination: Mol - Big

```{r}
                              ### Combination:  Mol - Big ###

#Behavior: Call Singing (Indirect Singing)
interaction_indirect_molbig <- glmmTMB(Singing ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                              family = nbinom2, data = lm_molbig)


additive_indirect_molbig <- glmmTMB(Singing ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                                 family = nbinom2, data = lm_molbig)

anova(interaction_indirect_molbig, additive_indirect_molbig)
```

```{r}
# Behavior: Courtship Singing (Direct Singing) 
interaction_direct_molbig <- glmmTMB(Active.Singing ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                            family = nbinom2, data = lm_molbig)


additive_direct_molbig <- glmmTMB(Active.Singing ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                               family = nbinom2, data = lm_molbig)

anova(interaction_direct_molbig, additive_direct_molbig)
```

```{r}
#Behavior: Mounting Attempt (Harassment) 
interaction_harass_molbig <- glmmTMB(Harassment ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                            family = nbinom2, data = lm_molbig)

additive_harass_molbig <- glmmTMB(Harassment ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                               family = nbinom2, data = lm_molbig)

anova(interaction_harass_molbig, additive_harass_molbig)

```

Then we analyze again for female behaviors per combination for the simplified models

## Combination: Bru - Big

```{r}
                                     ### Combination: Bru - Big ###
                                 
# Subset per behavior for this combination
resp_brubig <- resp %>% filter(Combination == "BruBru" | Combination == "BruBig"|
                            Combination == "BigBig" | Combination == "BigBru")

rej_brubig <- rej %>% filter(Combination == "BruBru" | Combination == "BruBig"|
                            Combination == "BigBig" | Combination == "BigBru")

#Behavior: Response
interaction_resp_brubig <- glmmTMB(cbind(Response, Response_fail) ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                              family = binomial, data = resp_brubig)

additive_resp_brubig <- glmmTMB(cbind(Response, Response_fail) ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                                 family = binomial, data = resp_brubig)

anova(interaction_resp_brubig, additive_resp_brubig)

```

```{r}
# Behavior: Rejection

interaction_rej_brubig <- glmmTMB(cbind(Hrejection, Rejection_fail) ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                              family = binomial, data = rej_brubig)

additive_rej_brubig <- glmmTMB(cbind(Hrejection, Rejection_fail) ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                                 family = binomial, data = rej_brubig)

anova(interaction_rej_brubig, additive_rej_brubig)

```

## Combination: Bru - Mol

```{r}
                                     ### Combination: Bru - Mol ###
# Subset per behavior for this combination
resp_brumol <- resp %>% filter(Combination == "BruBru" | Combination == "BruMol"|
                            Combination == "MolBru" | Combination == "MolMol")

rej_brumol <- rej %>% filter(Combination == "BruBru" | Combination == "BruMol"|
                            Combination == "MolBru" | Combination == "MolMol")
#Behavior: Response
interaction_resp_brumol <- glmmTMB(cbind(Response, Response_fail) ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                              family = binomial, data = resp_brumol)

additive_resp_brumol <- glmmTMB(cbind(Response, Response_fail) ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                                 family = binomial, data = resp_brumol)

anova(interaction_resp_brumol, additive_resp_brumol)

```

```{r}

# Behavior: Rejection

interaction_rej_brumol <- glmmTMB(cbind(Hrejection, Rejection_fail) ~ Male*Female + Type.F + Type.M + (1|Cage.males) + (1|Cage.females), 
                              family = binomial, data = rej_brumol)

additive_rej_brumol <- glmmTMB(cbind(Hrejection, Rejection_fail) ~ Male + Female + Type.F + Type.M + (1|Cage.males) + (1|Cage.females),
                                 family = binomial, data = rej_brumol)

anova(interaction_rej_brumol, additive_rej_brumol)

```

```{r}
# Behavior: Rejection w/ beta binomial distribution
## DHARMa shows overdispersion and zero-inflation, so we fit with beta binomial distribution to see if it improves the fit

interaction_rej_brumol_beta <- glmmTMB(cbind(Hrejection, Rejection_fail) ~ Male*Female + Type.F + Type.M + (1|Cage.males) + (1|Cage.females), 
                              family = betabinomial, data = rej_brumol)

additive_rej_brumol_beta <- glmmTMB(cbind(Hrejection, Rejection_fail) ~ Male + Female + Type.F + Type.M + (1|Cage.males) + (1|Cage.females),
                                 family = betabinomial, data = rej_brumol)

anova(interaction_rej_brumol_beta, additive_rej_brumol_beta)

## check again with DHARMa
# calculate the residuals for the model
m28_beta <- simulateResiduals(fittedModel = interaction_rej_brumol_beta, plot = F)

residuals(m28_beta)

#plot simulation output
plot(m28_beta)

#tests if the overall distribution conforms to expectations.
testUniformity(m28_beta)
testDispersion(m28_beta)
testZeroInflation(m28_beta)
testResiduals(m28_beta)

```

## Combination: Mol - Big

```{r}
                                     ### Combination: Mol - Big ###
 
                                      
# Subset per behavior for this combination
resp_molbig <- resp %>% filter(Combination == "BigBig" | Combination == "BigMol"|
                            Combination == "MolBig" | Combination == "MolMol")

rej_molbig <- rej %>% filter(Combination == "BigBig" | Combination == "BigMol"|
                            Combination == "MolBig" | Combination == "MolMol")

#Behavior: Response  
interaction_resp_molbig <- glmmTMB(cbind(Response, Response_fail) ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                              family = binomial, data = resp_molbig)

additive_resp_molbig <- glmmTMB(cbind(Response, Response_fail) ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                                 family = binomial, data = resp_molbig)

anova(interaction_resp_molbig, additive_resp_molbig)

```

```{r}
# Behavior: Rejection

interaction_rej_molbig <- glmmTMB(cbind(Hrejection, Rejection_fail) ~ Male*Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ), 
                              family = binomial, data = rej_molbig)

additive_rej_molbig <- glmmTMB(cbind(Hrejection, Rejection_fail) ~ Male + Female + Type.F + Type.M + (1|Cage.males ) + (1|Cage.females ),
                                 family = binomial, data = rej_molbig)

anova(interaction_rej_molbig, additive_rej_molbig)

```

## Mating for all combinations

```{r}

# Behavior: Mating for all combos

## Bru-Big

interaction_mating_brubig <- glmmTMB(Mating ~ Male*Female + Type.F + Type.M + (1|Cage.males) + (1|Cage.females), 
                            family = binomial, data = lm_brubig)

additive_mating_brubig <- glmmTMB(Mating ~ Male + Female + Type.F + Type.M + (1|Cage.males) + (1|Cage.females), 
                            family = binomial, data = lm_brubig)

anova(interaction_mating_brubig, additive_mating_brubig)


# GLM

## Solving convergence issues:

glm_int_brubig <- glm(Mating ~ Male*Female + Cage.males + Cage.females + Type.F + Type.M,  family = binomial, data = lm_brubig) 

glm_add_brubig <- glm(Mating ~ Male + Female + Cage.males + Cage.females + Type.F + Type.M,  family = binomial, data = lm_brubig) 

anova(glm(Mating ~ Male*Female + Cage.males + Cage.females + Type.F + Type.M,  family = binomial, data = lm_brubig), 
      glm(Mating ~ Male + Female + Cage.males + Cage.females + Type.F + Type.M,  family = binomial, data = lm_brubig),
      test="LRT")

# calculate AIC for GLM models
AIC(glm_int_brubig)
AIC(glm_add_brubig)


```

```{r}
## Bru-Mol
interaction_mating_brumol <- glmmTMB(Mating ~ Male*Female + Type.F + Type.M + (1|Cage.males) + (1|Cage.females), 
                            family = binomial, data = lm_brumol)

additive_mating_brumol <- glmmTMB(Mating ~ Male + Female + Type.F + Type.M + (1|Cage.males) + (1|Cage.females), 
                            family = binomial, data = lm_brumol)

anova(interaction_mating_brumol, additive_mating_brumol)

```

```{r}
## Mol-Big

interaction_mating_molbig <- glmmTMB(Mating ~ Male*Female + Type.F + Type.M + (1|Cage.males) + (1|Cage.females), 
                            family = binomial, data = lm_molbig)

additive_mating_molbig <- glmmTMB(Mating ~ Male + Female + Type.F + Type.M + (1|Cage.males) + (1|Cage.females), 
                            family = binomial, data = lm_molbig)

anova(interaction_mating_molbig, additive_mating_molbig)

# GLM

## Solving convergence issues:

glm_int_molbig <- glm(Mating ~ Male*Female + Cage.males + Cage.females + Type.F + Type.M,  family = binomial, data = lm_molbig) 

glm_add_molbig <- glm(Mating ~ Male + Female + Cage.males + Cage.females + Type.F + Type.M,  family = binomial, data = lm_molbig) 

anova(glm(Mating ~ Male*Female + Cage.males + Cage.females + Type.F + Type.M,  family = binomial, data = lm_molbig), 
      glm(Mating ~ Male + Female + Cage.males + Cage.females + Type.F + Type.M,  family = binomial, data = lm_molbig),
      test="LRT")

# calculate AIC for GLM models
AIC(glm_int_molbig )
AIC(glm_add_molbig)

```

# Goodness-of-fit & residual plots

We use the DHARMa package to simulate, calculate and plot residuals.

### Bru-Big combination

```{r}
# For each behavior, we test the best model for residuals, uniformity, overdispersion, and inflation
# calculate the residuals for the model
m1 <- simulateResiduals(fittedModel = additive_indirect_brubig, plot = F)

residuals(m1)

#plot simulation output
plot(m1)
  
#tests if the overall distribution conforms to expectations.
testUniformity(m1)
testDispersion(m1)
testZeroInflation(m1)
testResiduals(m1)
```

```{r}
# calculate the residuals for the model
m2 <- simulateResiduals(fittedModel = interaction_indirect_brubig, plot = F)

residuals(m2)

#plot simulation output
plot(m2)
  
#tests if the overall distribution conforms to expectations.
testUniformity(m2)
testDispersion(m2)
testZeroInflation(m2)
testResiduals(m2)
```

```{r}

# calculate the residuals for the model
m7 <- simulateResiduals(fittedModel = additive_direct_brubig, plot = F)

residuals(m7)

#plot simulation output
plot(m7)

#tests if the overall distribution conforms to expectations.
testUniformity(m7)
testDispersion(m7)
testZeroInflation(m7)
testResiduals(m7)
```

```{r}

# calculate the residuals for the model
m8 <- simulateResiduals(fittedModel = interaction_direct_brubig, plot = F)

residuals(m8)

#plot simulation output
plot(m8)

#tests if the overall distribution conforms to expectations.
testUniformity(m8)
testDispersion(m8)
testZeroInflation(m8)
testResiduals(m8)
```

```{r}

# calculate the residuals for the model
m19 <- simulateResiduals(fittedModel = additive_harass_brubig, plot = F)

residuals(m19)

#plot simulation output
plot(m19)

#tests if the overall distribution conforms to expectations.
testUniformity(m19)
testDispersion(m19)
testZeroInflation(m19)
testResiduals(m19)
```

```{r}

# calculate the residuals for the model
m20 <- simulateResiduals(fittedModel = interaction_harass_brubig, plot = F)

residuals(m20)

#plot simulation output
plot(m20)

#tests if the overall distribution conforms to expectations.
testUniformity(m20)
testDispersion(m20)
testZeroInflation(m20)
testResiduals(m20)
```

```{r}

# calculate the residuals for the model
m13 <- simulateResiduals(fittedModel = additive_resp_brubig, plot = F)

residuals(m13)

#plot simulation output
plot(m13)

#tests if the overall distribution conforms to expectations.
testUniformity(m13)
testDispersion(m13)
testZeroInflation(m13)
testResiduals(m13)
```

```{r}

# calculate the residuals for the model
m14 <- simulateResiduals(fittedModel = interaction_resp_brubig, plot = F)

residuals(m14)

#plot simulation output
plot(m14)

#tests if the overall distribution conforms to expectations.
testUniformity(m14)
testDispersion(m14)
testZeroInflation(m14)
testResiduals(m14)
```

```{r}

# calculate the residuals for the model
m25 <- simulateResiduals(fittedModel = additive_rej_brubig, plot = F)

residuals(m25)

#plot simulation output
plot(m25)

#tests if the overall distribution conforms to expectations.
testUniformity(m25)
testDispersion(m25)
testZeroInflation(m25)
testResiduals(m25)
```

```{r}

# calculate the residuals for the model
m26 <- simulateResiduals(fittedModel = interaction_rej_brubig, plot = F)

residuals(m26)

#plot simulation output
plot(m26)

#tests if the overall distribution conforms to expectations.
testUniformity(m26)
testDispersion(m26)
testZeroInflation(m26)
testResiduals(m26)
```

```{r}

# calculate the residuals for the model
m31 <- simulateResiduals(fittedModel = glm_int_brubig, plot = F)

residuals(m31)

#plot simulation output
plot(m31)

#tests if the overall distribution conforms to expectations.
testUniformity(m31)
testDispersion(m31)
testZeroInflation(m31)
testResiduals(m31)
```

```{r}

# calculate the residuals for the model
m32 <- simulateResiduals(fittedModel = glm_add_brubig, plot = F)

residuals(m32)

#plot simulation output
plot(m32)

#tests if the overall distribution conforms to expectations.
testUniformity(m32)
testDispersion(m32)
testZeroInflation(m32)
testResiduals(m32)
```

### Bru-Mol combination

```{r}
# For each behavior, we test the best model for residuals, uniformity, overdispersion, and inflation

# calculate the residuals for the model
m3 <- simulateResiduals(fittedModel = additive_indirect_brumol, plot = F)

residuals(m3)

#plot simulation output
plot(m3)

#tests if the overall distribution conforms to expectations.
testUniformity(m3)
testDispersion(m3)
testZeroInflation(m3)
testResiduals(m3)
```

```{r}
# For each behavior, we test the best model for residuals, uniformity, overdispersion, and inflation

# calculate the residuals for the model
m4 <- simulateResiduals(fittedModel = interaction_indirect_brumol, plot = F)

residuals(m4)

#plot simulation output
plot(m4)

#tests if the overall distribution conforms to expectations.
testUniformity(m4)
testDispersion(m4)
testZeroInflation(m4)
testResiduals(m4)
```

```{r}

# calculate the residuals for the model
m9 <- simulateResiduals(fittedModel = additive_direct_brumol, plot = F)

residuals(m9)

#plot simulation output
plot(m9)

#tests if the overall distribution conforms to expectations.
testUniformity(m9)
testDispersion(m9)
testZeroInflation(m9)
testResiduals(m9)
```

```{r}

# calculate the residuals for the model
m10 <- simulateResiduals(fittedModel = interaction_direct_brumol, plot = F)

residuals(m10)

#plot simulation output
plot(m10)

#tests if the overall distribution conforms to expectations.
testUniformity(m10)
testDispersion(m10)
testZeroInflation(m10)
testResiduals(m10)
```

```{r}

# calculate the residuals for the model
m21 <- simulateResiduals(fittedModel = additive_harass_brumol, plot = F)

residuals(m21)

#plot simulation output
plot(m21)

#tests if the overall distribution conforms to expectations.
testUniformity(m21)
testDispersion(m21)
testZeroInflation(m21)
testResiduals(m21)
```

```{r}

# calculate the residuals for the model
m22 <- simulateResiduals(fittedModel = interaction_harass_brumol, plot = F)

residuals(m22)

#plot simulation output
plot(m22)

#tests if the overall distribution conforms to expectations.
testUniformity(m22)
testDispersion(m22)
testZeroInflation(m22)
testResiduals(m22)
```

```{r}

# calculate the residuals for the model
m15 <- simulateResiduals(fittedModel = additive_resp_brumol, plot = F)

residuals(m15)

#plot simulation output
plot(m15)

#tests if the overall distribution conforms to expectations.
testUniformity(m15)
testDispersion(m15)
testZeroInflation(m15)
testResiduals(m15)
```

```{r}

# calculate the residuals for the model
m16 <- simulateResiduals(fittedModel = interaction_resp_brumol, plot = F)

residuals(m16)

#plot simulation output
plot(m16)

#tests if the overall distribution conforms to expectations.
testUniformity(m16)
testDispersion(m16)
testZeroInflation(m16)
testResiduals(m16)
```

```{r}

# calculate the residuals for the model
m27 <- simulateResiduals(fittedModel = additive_rej_brumol, plot = F)

residuals(m27)

#plot simulation output
plot(m27)

#tests if the overall distribution conforms to expectations.
testUniformity(m27)
testDispersion(m27)
testZeroInflation(m27)
testResiduals(m27)
```

```{r}

# calculate the residuals for the model
m28 <- simulateResiduals(fittedModel = interaction_rej_brumol, plot = F)

residuals(m28)

#plot simulation output
plot(m28)

#tests if the overall distribution conforms to expectations.
testUniformity(m28)
testDispersion(m28)
testZeroInflation(m28)
testResiduals(m28)
```

```{r}

# calculate the residuals for the model
m33 <- simulateResiduals(fittedModel = additive_mating_brumol, plot = F)

residuals(m33)

#plot simulation output
plot(m33)

#tests if the overall distribution conforms to expectations.
testUniformity(m33)
testDispersion(m33)
testZeroInflation(m33)
testResiduals(m33)
```

```{r}

# calculate the residuals for the model
m34 <- simulateResiduals(fittedModel = interaction_mating_brumol, plot = F)

residuals(m34)

#plot simulation output
plot(m34)

#tests if the overall distribution conforms to expectations.
testUniformity(m34)
testDispersion(m34)
testZeroInflation(m34)
testResiduals(m34)
```

### Mol-Big combination

```{r}
# calculate the residuals for the model
m5 <- simulateResiduals(fittedModel = additive_indirect_molbig, plot = F)

residuals(m5)

#plot simulation output
plot(m5)

#tests if the overall distribution conforms to expectations.
testUniformity(m5)
testDispersion(m5)
testZeroInflation(m5)
testResiduals(m5)
```

```{r}
# calculate the residuals for the model
m6 <- simulateResiduals(fittedModel = interaction_indirect_molbig, plot = F)

residuals(m6)

#plot simulation output
plot(m6)

#tests if the overall distribution conforms to expectations.
testUniformity(m6)
testDispersion(m6)
testZeroInflation(m6)
testResiduals(m6)
```

```{r}

# calculate the residuals for the model
m11 <- simulateResiduals(fittedModel = additive_direct_molbig, plot = F)

residuals(m11)

#plot simulation output
plot(m11)

#tests if the overall distribution conforms to expectations.
testUniformity(m11)
testDispersion(m11)
testZeroInflation(m11)
testResiduals(m11)
```

```{r}

# calculate the residuals for the model
m12 <- simulateResiduals(fittedModel = interaction_direct_molbig, plot = F)

residuals(m12)

#plot simulation output
plot(m12)

#tests if the overall distribution conforms to expectations.
testUniformity(m12)
testDispersion(m12)
testZeroInflation(m12)
testResiduals(m12)
```

```{r}

# calculate the residuals for the model
m23 <- simulateResiduals(fittedModel = additive_harass_molbig, plot = F)

residuals(m23)

#plot simulation output
plot(m23)

#tests if the overall distribution conforms to expectations.
testUniformity(m23)
testDispersion(m23)
testZeroInflation(m23)
testResiduals(m23)
```

```{r}

# calculate the residuals for the model
m24 <- simulateResiduals(fittedModel = interaction_harass_molbig, plot = F)

residuals(m24)

#plot simulation output
plot(m24)

#tests if the overall distribution conforms to expectations.
testUniformity(m24)
testDispersion(m24)
testZeroInflation(m24)
testResiduals(m24)
```

```{r}

# calculate the residuals for the model
m17 <- simulateResiduals(fittedModel = additive_resp_molbig, plot = F)

residuals(m17)

#plot simulation output
plot(m17)

#tests if the overall distribution conforms to expectations.
testUniformity(m17)
testDispersion(m17)
testZeroInflation(m17)
testResiduals(m17)

```

```{r}

# calculate the residuals for the model
m18 <- simulateResiduals(fittedModel = interaction_resp_molbig, plot = F)

residuals(m18)

#plot simulation output
plot(m18)

#tests if the overall distribution conforms to expectations.
testUniformity(m18)
testDispersion(m18)
testZeroInflation(m18)
testResiduals(m18)

```

```{r}
# calculate the residuals for the model
m30 <- simulateResiduals(fittedModel = interaction_rej_molbig, plot = F)

residuals(m30)

#plot simulation output
plot(m30)

#tests if the overall distribution conforms to expectations.
testUniformity(m30)
testDispersion(m30)
testZeroInflation(m30)
testResiduals(m30)
```

```{r}

## !!!! POTENTIAL ISSUES WITH ZERO INFLATION

# calculate the residuals for the model
m30 <- simulateResiduals(fittedModel = interaction_rej_molbig, plot = F)

residuals(m30)

#plot simulation output
plot(m30)

#tests if the overall distribution conforms to expectations.
testUniformity(m30)
testDispersion(m30)
testZeroInflation(m30)
testResiduals(m30)
```

```{r}

# calculate the residuals for the model
m35 <- simulateResiduals(fittedModel = glm_int_molbig, plot = F)

residuals(m35)

#plot simulation output
plot(m35)

#tests if the overall distribution conforms to expectations.
testUniformity(m35)
testDispersion(m35)
testZeroInflation(m35)
testResiduals(m35)
```

```{r}

# calculate the residuals for the model
m36 <- simulateResiduals(fittedModel = glm_add_molbig, plot = F)

residuals(m36)

#plot simulation output
plot(m36)

#tests if the overall distribution conforms to expectations.
testUniformity(m36)
testDispersion(m36)
testZeroInflation(m36)
testResiduals(m36)
```
