---
title: "Calculating the Strength of Reproductive Isolation"
output: pdf_document
date: "2026-09-06"
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Strength of RI

Here, we calculate the strength of RI using the formula from [@sobel2014], where strength of RI is: 1 - 2\*(heterospecific/ (heterospecific + conspecific)). We use the simple bootstrapping method described by [@zhang2024role].

Changes from v1 to v2: Calculation of _female acceptance_ instead of _female rejection_ by calculating acceptance as 1-female rejection. 

v2 also calculates the absolute and relative contributions of each behavior to total reproductive isolation as described by [@ramsey2003components].

### Packages and libraries

```{r}
#set wd
setwd('/dss/dsslegfs01/pr53da/pr53da-dss-0029/big_assortative')

# load libraries and packages
library(ggplot2)
library(dplyr)
library(tidyr)

#Load the file. 
hoppers <- read.csv("hoppers2_femresp_corrected.csv", sep = ";")

#Creating a fail column to convert into sucess/fail data 
hoppers$Response_fail <- hoppers$Total.Singing - hoppers$Response
hoppers$Rejection_fail <- hoppers$Harassment - hoppers$Hrejection

# Ensure origin variables are factors
hoppers$Type.F <- factor(hoppers$Type.F)  # lab / wild
hoppers$Type.M <- factor(hoppers$Type.M)

# Flags and combined origin
hoppers$FemaleWild <- hoppers$Type.F == "wild"
hoppers$MaleWild   <- hoppers$Type.M == "wild"
hoppers$AnyWild    <- hoppers$FemaleWild | hoppers$MaleWild
hoppers$BothLab    <- !hoppers$FemaleWild & !hoppers$MaleWild

hoppers$OriginPair <- interaction(hoppers$Type.F, hoppers$Type.M, sep = "_")

#Response 
hoppers$ResponseNorm <- (hoppers$Response/hoppers$Total.Singing)  
resp <- hoppers[-which(hoppers$ResponseNorm == "NaN"), ]      
#resp <- resp[-which(resp$ResponseNorm == "Inf"), ]

#Rejection
hoppers$RejectNorm <- (hoppers$Hrejection/hoppers$Harassment)
rej <- hoppers[-which(hoppers$RejectNorm == "NaN"), ] 

# Response: remove rows where Total.Singing = 0 (NaN in proportion)
resp <- subset(hoppers, !is.nan(ResponseNorm))

# Rejection: remove rows where Harassment = 0 (NaN in proportion)
rej  <- subset(hoppers, !is.nan(RejectNorm))

# add before the behavior_list
hoppers$AcceptNorm <- 1 - hoppers$RejectNorm 

```

## All combinations

```{r}
# define the species pairs
species_pairs <- list( c("Big", "Bru"), c("Big", "Mol"), c("Bru", "Mol"))
```

### Behaviors list

```{r}
## list all behaviors and data sets, here we specify the type of distribution for the parametric bootstrapping.
## we also add a sign for each behavior, as female rejection gives a negative RI value (more heterospecific interactions than conspecific), and we want it to be to a positive value for comparing it with other values

##v2: sign for each behavior is no longer relevant, as we are now calculating female acceptance instead of female rejection

behavior_list <- list(
  Mounting_Attempt = list(
    data = hoppers,
    behavior_col = "Harassment",
    dist         = "nbinom",
    sign         = +1
  ),
  Call_Singing = list(
    data = hoppers,
    behavior_col = "Singing",
    dist         = "nbinom",
    sign         = +1
  ),
  Courtship_Singing = list(
    data = hoppers,
    behavior_col = "Active.Singing",
    dist         = "nbinom",
    sign         = +1
  ),
  Female_Response = list(
    data = hoppers,
    behavior_col = "ResponseNorm",
    dist         = "binomial",
    sign         = +1
  ),
   #Female_Rejection = list(
    #data = hoppers,
    #behavior_col = "RejectNorm",
    #dist         = "binomial",
    #sign         = -1
  #),
  Female_Acceptance = list(
    data = hoppers,
    behavior_col = "AcceptNorm",
    dist         = "binomial",
    sign         = +1    
  ),
  Mating = list(
    data = hoppers,
    behavior_col = "Mating",
   dist         = "binomial",
    sign         = +1
  )
)
```

## Computing RI

```{r}
# Calculate RI for each pair for each behavior
compute_RI_pair <- function(df, sp1, sp2, male_col, female_col, behavior_col) {
  cons_rows <- 
    (df[[male_col]] == sp1 & df[[female_col]] == sp1) |
    (df[[male_col]] == sp2 & df[[female_col]] == sp2)
  
  het_rows  <- 
    (df[[male_col]] == sp1 & df[[female_col]] == sp2) |
    (df[[male_col]] == sp2 & df[[female_col]] == sp1)
  
  cons_vals <- df[[behavior_col]][cons_rows]
  het_vals  <- df[[behavior_col]][het_rows]
  
  C <- mean(cons_vals, na.rm = TRUE)
  H <- mean(het_vals,  na.rm = TRUE)
  
  RI <- 1 - 2 * (H / (H + C))
  
  list(
    C = C, H = H, RI = RI,
    cons_vals = cons_vals,
    het_vals  = het_vals,
    cons_idx  = cons_rows,
    het_idx   = het_rows
  )
}

```

## Simple bootstrapping

```{r}
## Bootstrap function for 1 species pair & 1 behavior
## This resamples rows WITH replacement, maintaining the species pair structure.

bootstrap_RI_simple <- function(df, sp1, sp2, male_col, female_col, behavior_col,
                                sign = 1,
                                nboot = 10000) {
  
  base <- compute_RI_pair(df, sp1, sp2, male_col, female_col, behavior_col)
  
  cons_vals <- base$cons_vals[!is.na(base$cons_vals)]
  het_vals  <- base$het_vals[!is.na(base$het_vals)]
  
  n_con <- length(cons_vals)
  n_het <- length(het_vals)
  
  if (n_con == 0 || n_het == 0) {
    stop("No conspecific or heterospecific data for this species pair.")
  }
  
  RI_boot <- numeric(nboot)
  
  for (b in seq_len(nboot)) {
    sim_con <- sample(cons_vals, n_con, replace = TRUE)
    sim_het <- sample(het_vals,  n_het, replace = TRUE)
    
    Cb <- mean(sim_con)
    Hb <- mean(sim_het)
    
    RIb <- 1 - 2 * (Hb / (Hb + Cb))
    RI_boot[b] <- sign * RIb   # apply sign flip inside bootstrap -- can be removed
  }
  
  # Signed final values
  RI_signed <- sign * base$RI
  
  CI <- quantile(RI_boot, c(0.025, 0.975), na.rm = TRUE)
  
  list(
    C        = base$C,
    H        = base$H,
    RI       = RI_signed,
    CI_lower = CI[1],
    CI_upper = CI[2],
    n_con    = n_con,
    n_het    = n_het,
    RI_boot  = RI_boot
  )
}

```

```{r}
## Loop over all behaviors × species pairs with 10000 bootstraps

results_simple <- data.frame()

for (beh in names(behavior_list)) {
  
  df    <- behavior_list[[beh]]$data
  bcol  <- behavior_list[[beh]]$behavior_col
  dist  <- behavior_list[[beh]]$dist
  sign  <- ifelse(is.null(behavior_list[[beh]]$sign), 1, behavior_list[[beh]]$sign)
  
  for (i in seq_along(species_pairs)) {
    
    sp1 <- species_pairs[[i]][1]
    sp2 <- species_pairs[[i]][2]
    
    base <- compute_RI_pair(df, sp1, sp2, "Male", "Female", bcol)
    
    n_con <- sum(!is.na(base$cons_vals))
    n_het <- sum(!is.na(base$het_vals))
    
    if (n_con == 0 || n_het == 0) next
    
    out <- bootstrap_RI_simple(
      df           = df,
      sp1          = sp1,
      sp2          = sp2,
      male_col     = "Male",
      female_col   = "Female",
      behavior_col = bcol,
      sign         = sign,
      nboot        = 10000
    )
    
    row <- data.frame(
      behavior     = beh,
      species_pair = paste(sp1, sp2, sep = "-"),
      C            = out$C,
      H            = out$H,
      RI           = out$RI,
      CI_lower     = out$CI_lower,
      CI_upper     = out$CI_upper,
      n_cons       = out$n_con,
      n_het        = out$n_het
    )
    
    results_simple <- rbind(results_simple, row)
  }
}
results_simple
```

```{r}
## Plotting results
plot_RI_single <- function(results_df,
                           dodge_width = 0.6,
                           point_size = 3,
                           line_size = 0.7) {
  
  library(ggplot2)
  
  # Desired behavior order (adjust names if needed)
  desired_order <- c("Call_Singing",
                     "Courtship_Singing",
                     "Female_Response",
                     "Mounting_Attempt",
                     "Female_Acceptance",
                     "Mating")
  
  # Ensure behavior column respects this order
  present_beh <- intersect(desired_order, unique(as.character(results_df$behavior)))
  results_df$behavior <- factor(results_df$behavior, levels = present_beh)
  
  # Factor species_pair for consistent coloring
  results_df$species_pair <- factor(results_df$species_pair)
  
  # Y-axis breaks every 0.2 from 0 to 1
  ybreaks <- seq(-0.4, 1, by = 0.2)
  
  # Plot
  p <- ggplot(results_df, aes(x = behavior, y = RI, color = species_pair)) +
    geom_point(position = position_dodge(width = dodge_width),
               size = point_size) +
    geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper),
                  position = position_dodge(width = dodge_width),
                  width = 0.2,
                  linewidth = line_size) +
    scale_y_continuous(breaks = ybreaks, limits = c(-0.4,1)) +
    theme_bw(base_size = 14) +
    labs(
      x = "Behavior",
      y = "Reproductive isolation (RI)",
      color = "Species pair"
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid.major.x = element_blank()
    )
  
  return(p)
}

```

```{r}
## plot

#p_bootsimple <- 
simple_plot <- plot_RI_single(results_simple)
simple_plot
```

```{r}
## Save & export plots

p_simple <- plot_RI_single(results_simple)
ggsave("RI_simple_10k_v2.svg", p_simple, device="svg", width=10, height=6, units="in")
ggsave("RI_simple_10k_v2.png", p_simple, device="png", width=10, height=6, units="in", dpi=500)
```

## Calculation of absolute and relative contributions

```{r}
# Behaviors in their fixed temporal courtship order.
behavior_sequence <- list(
  list(name = "Call_Singing",       col = "Singing",        sign = +1),
  list(name = "Courtship_Singing",  col = "Active.Singing",  sign = +1),
  list(name = "Female_Response",    col = "ResponseNorm",    sign = +1),
  list(name = "Mounting_Attempt",   col = "Harassment",      sign = +1),
  list(name = "Female_Acceptance",  col = "AcceptNorm",      sign = +1),
  list(name = "Mating",             col = "Mating",          sign = +1)
)

bootstrap_AC_sequence <- function(df, sp1, sp2, male_col, female_col,
                                   behavior_sequence, nboot = 10000) {

  cons_rows_idx <- which((df[[male_col]] == sp1 & df[[female_col]] == sp1) |
                          (df[[male_col]] == sp2 & df[[female_col]] == sp2))
  het_rows_idx  <- which((df[[male_col]] == sp1 & df[[female_col]] == sp2) |
                          (df[[male_col]] == sp2 & df[[female_col]] == sp1))

  n_con <- length(cons_rows_idx)
  n_het <- length(het_rows_idx)
  if (n_con == 0 || n_het == 0) stop("No conspecific or heterospecific data for this pair.")

  n_beh <- length(behavior_sequence)
  beh_names <- sapply(behavior_sequence, function(x) x$name)

  # storage: one row per bootstrap replicate, one column per behavior
  RI_mat <- matrix(NA_real_, nrow = nboot, ncol = n_beh, dimnames = list(NULL, beh_names))
  AC_mat <- matrix(NA_real_, nrow = nboot, ncol = n_beh, dimnames = list(NULL, beh_names))

  for (b in seq_len(nboot)) {

    # resample TRIALS ONCE per replicate — same resampled rows feed every behavior
    sim_con_idx <- sample(cons_rows_idx, n_con, replace = TRUE)
    sim_het_idx <- sample(het_rows_idx,  n_het, replace = TRUE)

    sim_con <- df[sim_con_idx, ]
    sim_het <- df[sim_het_idx, ]

    RI_row <- numeric(n_beh)
    for (k in seq_len(n_beh)) {
      col  <- behavior_sequence[[k]]$col
      sign <- behavior_sequence[[k]]$sign

      C <- mean(sim_con[[col]], na.rm = TRUE)
      H <- mean(sim_het[[col]], na.rm = TRUE)

      RI_row[k] <- sign * (1 - 2 * (H / (H + C)))
    }
    RI_mat[b, ] <- RI_row

    # sequential absolute contribution: each behavior only acts on the
    # fraction of heterospecific pairs not already excluded upstream
    cum <- 0
    AC_row <- numeric(n_beh)
    for (k in seq_len(n_beh)) {
      AC_row[k] <- RI_row[k] * (1 - cum)
      cum <- cum + AC_row[k]
    }
    AC_mat[b, ] <- AC_row
  }

  # relative contribution per replicate: AC_i / total AC (row-wise)
  total_AC <- rowSums(AC_mat)
  RC_mat <- AC_mat / total_AC

  summarize_col <- function(mat) {
    data.frame(
      behavior = beh_names,
      mean     = colMeans(mat, na.rm = TRUE),
      CI_lower = apply(mat, 2, quantile, probs = 0.025, na.rm = TRUE),
      CI_upper = apply(mat, 2, quantile, probs = 0.975, na.rm = TRUE)
    )
  }

  list(
    RI_summary = summarize_col(RI_mat),
    AC_summary = summarize_col(AC_mat),
    RC_summary = summarize_col(RC_mat),
    RI_mat = RI_mat, AC_mat = AC_mat, RC_mat = RC_mat
  )
}

```

```{r}
#Loop it over all three species pairs
AC_results <- list()

for (i in seq_along(species_pairs)) {
  sp1 <- species_pairs[[i]][1]
  sp2 <- species_pairs[[i]][2]
  pair_name <- paste(sp1, sp2, sep = "-")

  AC_results[[pair_name]] <- bootstrap_AC_sequence(
    df = hoppers, sp1 = sp1, sp2 = sp2,
    male_col = "Male", female_col = "Female",
    behavior_sequence = behavior_sequence,
    nboot = 10000
  )
}

# assemble a tidy table for the manuscript
AC_table <- do.call(rbind, lapply(names(AC_results), function(p) {
  df <- AC_results[[p]]$AC_summary
  df$species_pair <- p
  df
}))
RC_table <- do.call(rbind, lapply(names(AC_results), function(p) {
  df <- AC_results[[p]]$RC_summary
  df$species_pair <- p
  df
}))
```

```{r}
#export tidy tables
write.table(AC_table, file = "AC_results.csv", col.names = TRUE)
write.table(RC_table, file = "RC_results.csv", col.names = TRUE)

```

## References

